\textit{This chapter is based on a joint research with Cecilia García-Peñalosa (Aix-Marseille University) and Tanguy van Ypersele (Aix-Marseille University).}\\
\vspace{1em}\\
\noindent \textbf{Abstract}: The increase in employment polarization observed in a number of high-income economies has coincided with a reduction in inter-generational mobility. This paper uses data for two British cohorts that entered the labour market at two points in time that differed considerably in terms of the structure of employment to re-examine the drivers of mobility. 
We differ from the existing literature in two aspects. First, we focus on employment categories rather than income, thus obtaining dynamics that can be understood in terms of changes in the structure of employment. Second, we argue that understanding inter-generational dynamics requires considering how individuals move from their entry jobs into other employment categories, i.e. understanding intra-generational mobility. 
The data indicate that occupational changes over the individual’s career are an important source of mobility, with large shares of those in low-paying (respectively, middling) occupations moving into middling (resp. high-paying) ones. When we compare the two cohorts we find that these two sources of mobility have declined for the younger cohort and that, whatever the initial occupation, parental income has become more important in leading to occupational upgrading. 
Moreover, the impact of parental income increased the most in the regions where the share of middling employment fell the most, suggesting that increased employment polarization may be one of the factors behind the observed decline in mobility.\\
\vspace{1em}\\
\noindent\textbf{Keywords:} Inter-generational mobility, Job polarization.\\
\noindent\textbf{JEL Codes:} J62, J21, J24.

\clearpage
\chaptertoc{}

\pagebreak

\begin{refsection}
    
    \section{Introduction}\label{chap2-introduction}
    \input{chap2/chapters/01-introduction}
    
    \section{Theoretical framework} \label{chap2-model}
    \input{chap2/chapters/02-theory}
    
    \section{Data and employment polarization} \label{chap2-data}
    \input{chap2/chapters/03-data}
    
    \section{Empirical specification} \label{chap2-specification}
    \input{chap2/chapters/04-specification}
    
    \section{Patterns of mobility} \label{chap2-mobility}
    \input{chap2/chapters/05-mobility}
    
    \section{Mobility and polarization at the regional level}\label{chap2-regional}
    \input{chap2/chapters/06-regional}
    
    \section{Conclusion} \label{chap2-conclusion}
    \input{chap2/chapters/07-conclusion}
    
    \printbibliography[heading=subbibintoc]
    
    \clearpage
    \addsec{Appendices}
    \renewcommand{\thesubsection}{\thechapter.\Alph{subsection}}
    
    \subsection{Model derivations}\label{chap2-app-model}
    \input{chap2/chapters/A-model}
    \clearpage
    \subsection{Data and summary statistics}\label{chap2-app-data}
    \input{chap2/chapters/B-data}
    \clearpage
    \subsection{Logistic regressions}\label{chap2-app-logit}
    \input{chap2/chapters/C-logistic}
    \clearpage
    \subsection{Accounting for education} \label{chap2-app-education}
    \input{chap2/chapters/D-education}
    \clearpage
    \subsection{Additional material}\label{chap2-app-additional}
    \input{chap2/chapters/E-additional}
    \clearpage
    \subsection{Decomposing the effect of parental income}\label{chap2-app-decomposing}
    \input{chap2/chapters/F-decomposing}
    \clearpage
    \subsection{Robustness checks}\label{chap2-app-robustness}
    \input{chap2/chapters/G-robustness}

\end{refsection}