First of all, I am grateful to Cecilia García Peñalosa and Marc Sangnier. I would like to thank them for their time and help all along this journey. I once summarized those three very diversified chapters to someone, and that person asked me: ``Who's the madman supervising this?''. I could never thank them enough for their intellectual flexibility in that respect. I have learned so much from them since they both decided to supervise my master's dissertation, then this thesis. I would like to thank them for being two brilliant professors and wonderful human beings.

I wish to thank the other members of my Ph.D. committee for their time and very accurate comments. 
Yanos Zylberberg invited me to the University of Bristol and provided me with incredibly useful advice and comments during those few months. I am grateful for his humor and enthusiasm which helped me to cope with this job market year. 
Xavier Raurich was my professor when I was in Erasmus at the Universitat de Barcelona. I am grateful for his interest and support since then. I am deeply convinced that his course motivated me to enroll in this Ph.D. journey while I was only a master's student. 
Giovanni Pica and Thierry Verdier are among the scholars that I enjoyed the most reading their papers. I am grateful to them as they helped improve the quality of my thesis through their comments.

I am grateful to the Aix-Marseille School of Economics as an institution and to all its members that I met during my time there. Karine Gente and Carine Nourry allowed me to do an internship as a research assistant when I was only in my third year. I am grateful to them for this opportunity which was also key in the decision to enroll in a Ph.D. years after. 
I am also grateful to all professors that I met and discussed with along those years, especially, Patricia Augier, Nicolas Berman, Renaud Bourlès, Bruno Decreuse, Eva Moreno-Galbis, Lorenzo Rotunno, Avner Seror, Tanguy van Ypersele, and Roberta Ziparo. 
I wish to thank the administrative team for their support and effectiveness, especially, Marine Boléa, Patrice Cacciuttolo, Anne-Sophie Leclercq Therond, Elisabeth Lhuillier, and Bernadette Vouriot.
I would like to express my gratitude to all former and current PhD students from AMSE with whom I shared very enjoyable moments, especially, Daniela Arlia, Stéphane Benveniste, Laurène Bocognagno, Kenza Elass, Jordan Loper, Nandeeta Neerunjun, Julieta Peveri, Morgan Raux, Laura Sénécal, Rosnel Sessinou, Mathilde Valéro, Sarah Vincent, and Rémi Vives.

My deepest thanks are addressed to both my family and my friends for their support all along those years, especially, 
Jade Benzaki, Tamara Casanova, Hugo Chirossel, Nicolas Court,
Ginette Galliano, Marc Galliano, Marie-Anne Galliano, Nicolas Galliano, Pierette Galliano,
Ombeline Jullien de Pommerol, Philippe Lavall, Fabien Martinet, Luke Mcgee O'Collins, Lucas Motta,
Jérôme Petit, Maryse Petit, Maxime Petit, 
Julie Pierson, Céline Prella, Jean-Marc Prella,
and Yohann Serot.